\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-02-15 trunk/europa-basics.tex
rearranging text from concepts into europa-basics

@@

Constraint Satisfaction Programming (also known as Constraint
Programming (CP)) is a discipline that provides a generic framework
for representing, solving and making logical inference on
constraints. A complete treatment of this discipline is given in
\cite{marriott98,apt03} and very concise introductions are provided in
\cite{bartak99,lustig01}.

A constraint programming problem consists of a set of variables $V=
{x_1,..,x_N}$, where each variable takes values from a domain
$d_1,..,d_N$; in this chapter we will deal only with discrete and
finite domains. Given a defined conjunctive set of constraints on the
variables: $C=\{c_1(x_1,..,x_N), ..., c_K(x_1,..,x_N)\}$, the
objective is to find one or more value assignments to $V$ where all
constraints are satisfied. To solve a problem, CP techniques use
logical inference to perform Constraint Propagation, which consists of
two operations:

@@

A set of constraints and variables that describe a problem domain are
typically represented as a Constraint Network, where the variables are
nodes and constraints are arcs between the variables. Having this
network representation in mind, a pair of variables $x,y$ is said to
be arc consistent if for every value in $x$'s domain, there is a value
in $y$'s domain that is consistent with all the constraints that
connect $x$ and $y$.  A naive approach to ensuring arc consistency
consists of cycling through all the variable pairs and performing
constraint propagation until there are no variable domain changes. The
AC-3 algorithm \cite{mackworth77}, a popular implementation, improves
over the naive approach by ignoring constraints whose variables were
not affected since the last iteration.

Arc Consistency and other, more stringent consistency definitions can
be used to efficiently prove that a CP problem is
unsatisfiable. However, they generally cannot prove that a CP problem
is satisfiable and provide specific variable assignments that
constitute solutions. For that, search algorithms like global
backtracking \comment{citation needed} or local search are used; these
use consistency as an efficient way to prune the search space
\cite{cp06}. In theory, solving a CP problem is NP-Hard
\cite{ghallab04}, but often very efficient in practice using a number
of consistency and search algorithms that are well understood.

CP is usually implemented as part of a programming language and
constraints are usually represented as objects \cite{puget95}. Any
constraint that the user is able to reason about can be introduced
into the system and, as long as the constraint propagation protocols
specified by the host CP system are enforced, it will be
indistinguishable from any other "primitive'' CP constraint, such as
$\leq$ or $\geq$.

>>>>

\section{Foundational Concepts}
\label{sec:concepts}

\rx is an adaptive, artificial intelligence based controller and
provides general framework for building reasoning systems for
real-world autonomous vehicles. At MBARI \rx is used for AUV control;
another instantiation of the system is being used for control of a
terrestrial personal robot \cite{pr2, Meeussen:2010dn}. The
development of \rx has been targeted at surveying a number of
oceanographic features which are dynamic and unpredictable
spatio-temporally. Typically this requires our AUVs to balance the
goals of coverage spatially while opportunistically following features
of scientific interest and to do so while being operationally aware of
its own limitations in terms of resources (typically the battery state
of charge) and overall proximity to other observational assets for
obtaining scientific ground-truth. We are therefore interested in
representational frameworks that allow robots to pursue long-term
objectives while choosing short-term gain and can gracefully deal with
exogenous or endogenous change.

To enable this responsiveness to external observations, the agent has
to be able to synthesize plans insitu and to re-plan. We use a
temporal constraint-based planner \eu with a demonstrated legacy of
having flown on NASA space missions \cite{jonsson00,bresina05,
  barreiro09}. Our autonomy architecture brings three key innovations
for AUV adaptation: the use of flexible plan representations,
compositional control with the use of partitioned networks and
off-line learning to inform insitu environmental state estimation. In
this section we detail some of the key concepts relevant to flexible
plan representation. Section \ref{sec:arch} has details the
\rx partitioned architecture and estimation driven vehicle
adaptation.

\subsection{The basics}
\label{sec:basics}

\eu uses a \emph{domain model} written in a declarative language
(NDDL: New Domain Description Language), together with initial
conditions and goals also specified in NDDL, to construct a set of
temporal relations that must be true at the start time. These models
include assertions about the physics of the vehicle, i.e how it
responds to external stimulus and internally driven goals. By
propagating these relations forward using Simple Temporal Networks
\cite{dechter91} and applying goal constraints, \eu can select a set
of conditions that should be true in the future, where some of these
conditions will correspond to actions the agent must take. The planner
can backtrack and try another path during search if a goal cannot be
reached while being capable of discarding unachievable goals.

Traditionally robot execution has relied on dispatching commands at
precise times. Such linear sequences of precisely timed commands give
no ability to adjust execution on the basis of sensory
information. Although some commands can issue tests on sensor
readings, these tests have the objective of verifying whether expected
execution conditions are occuring. If not, the state of the system is
declared off-nominal and execution of the sequence of commands is
interrupted. More recently, executives have been proposed and
implemented that significantly broaden the way robots can be commanded
\cite{mus98,alami:1998p820}. For example, the Remote Agent executive
interpreted a \textit{temporally flexible plan} which represents each
start time as a variable and contains an explicit network of bounded
delay constraints between such variables.

\begin{figure}[!t]
\centering
\includegraphics[scale=0.35]{figs/flexible-timelines.pdf}
\caption{\small Tokens with flexible temporal intervals and parametric
  constraints between tokens. This example shows the triggering of a
  water sampler based on a feature threshold while the vehicle
  Yo-Yo's. The \texttt{Waypoint\_Yo-Yo} token has a flexible duration,
  start \& end times.}
\label{fig:flex-timelines}
\vskip+0.1cm
\end{figure}

Unlike a traditional fixed time-tagged command sequences, such
flexible plans leave room for adaptation at execution time. When the
executive considers when to start a task, it propagates information
through the constraint network, computes a time bound for the
variable, selects an actual execution time within the bound, and
starts the task at that time. Temporally flexible plans therefore,
express a \textit{range of possible outcomes} of the robots
interaction with the environment within which the executive can elect
at run time the most appropriate one for the actual execution
conditions. The fact that constraints are explicitly represented
ensures that through constraint propagation the executive will respect
global limits expressed in the plan (e.g., don't start a task until a
certain condition has been satisfied but still satisfying some global
deadline). Such flexibility is critical when dealing with dynamic
ocean conditions where precise timing of a robotic action might be
indeterminate. Further, the advantage of flexibility can be contrasted
with the consequences of the intrinsic inflexibiliy of traditional
command sequences. Because they are inflexible, sequences must
necessarily be designed considering worst case scenarios.

\eu uses a \emph{state variable} representation to describe the
evolution of state over time. The instantiated history of such state
variable evolution over a temporal horizon we call \emph{timelines}
and which represent a single thread in the execution of a concurrent
system. At any given time each thread can execute a single procedure.
Thus each timeline consists of a sequence of procedures which
encapsulate and describe state evolution; we call these instantiated
atomic entities \emph{tokens}.  A token therefore describes a
procedure invocation, the state variables on which it can occur, the
parameter values of the procedure, and the time values defining the
interval. We allow encapsulation of uncertainty within these tokens
with a range of start and end times and parameters, all of which are
encoded as variables. A constraint solver in turn manipulates these
variables defined in a \eu domain model. For example, consider
a scientific need to take a water samples 100 meters from a hotspot
while an AUV is performing a Yo-Yo in the water-column. Two samples
are needed if the feature's signal is above a threshold; one
otherwise. The token that is capturing the sensory threshold has a
parametric constraints to the token which fires the requisite water
sampler. In addition, the start time of the water sampling procedure
is highly dependant on the variability of sub-sea currents and actual
speed of the vehicle. Therefore a number of values are possible for
the start times and duration of the sampling all of which are valid
combinations for desired outcomes.

% \begin{figure}
% \centering
% \includegraphics[scale=0.3]{figs/Allen-algebra.pdf}
% \caption{\small Temporal relations defined within the planner are
%   based on \texttt{Allen Algebra} relations shown above.}
% \label{fig:allen-algebra}
% \vskip-0.3cm
% \end{figure}

@@

Constraint Satisfaction Programming (also known as Constraint
Programming (\textsf{CP})) is a discipline that provides a generic framework
for representing, solving and making logical inference on
constraints. A complete treatment of this discipline is given in
\cite{marriott98,apt03} and very concise introductions are provided in
\cite{bartak99,lustig01}.

A constraint programming problem consists of a set of variables $V=
{x_1,..,x_N}$, where each variable takes values from a domain
$d_1,..,d_N$; in this chapter we will deal only with discrete and
finite domains. Given a defined conjunctive set of constraints on the
variables: $C=\{c_1(x_1,..,x_N), ..., c_K(x_1,..,x_N)\}$, the
objective is to find one or more value assignments to $V$ where all
constraints are satisfied. To solve a problem, \textsf{CP} techniques use
logical inference to perform Constraint Propagation, which consists of
two operations:

@@

A set of constraints and variables that describe a problem domain are
typically represented as a Constraint Network, where the variables are
nodes and constraints are arcs between the variables. Having this
network representation in mind, a pair of variables $x,y$ is said to
be arc consistent if for every value in $x$'s domain, there is a value
in $y$'s domain that is consistent with all the constraints that
connect $x$ and $y$.  A naive approach to ensuring arc consistency
consists of cycling through all the variable pairs and performing
constraint propagation until there are no variable domain changes. The
AC-3 algorithm \cite{mackworth77}, a popular implementation, improves
over the naive approach by ignoring constraints whose variables were
not affected since the last iteration. \comment{A figure of a
  constraint network preferably of the shopping example would be
  useful here.}

Arc Consistency and other, more stringent consistency definitions can
be used to efficiently prove that a \textsf{CP} problem is
unsatisfiable. However, they generally cannot prove that a \textsf{CP}
problem is satisfiable and provide specific variable assignments that
constitute solutions. For that, search algorithms like global
backtracking \comment{citation needed} or local search are used; these
use consistency as an efficient way to prune the search space
\cite{cp06}. In theory, solving a \textsf{CP} problem is NP-Hard
\cite{ghallab04}, but often very efficient in practice using a number
of consistency and search algorithms that are well understood.

\textsf{CP} is usually implemented as part of a programming language
and constraints are usually represented as objects \cite{puget95}. Any
constraint that the user is able to reason about can be introduced
into the system and, as long as the constraint propagation protocols
specified by the host \textsf{CP} system are enforced, it will be
indistinguishable from any other "primitive'' \textsf{CP} constraint,
such as $\leq$ or $\geq$.
